\chapter{\textsc{Typesetter's Notes}}



My goal was to present, as best I could, Dr. Lush's vision for \textit{Animal
Breeding Plans} without imposition of my own ideas. I've limited my role
strictly to that of typesetter. I haven't corrected any spelling or
typographical errors (e.g., missing hyphens), and the diligent reader will
note that references are formatted differently in different chapters. I didn't
feel it was appropriate to make such changes in this edition.
There are some minor differences between old and new that were unavoidable. For
example, due to differences in margins and typeface sizes, page numbers in this
new edition do not match those in the original. Table layouts are slightly
different even if they're identical in content. The exception to this rule is
Table 22, which I divided into two separate tables. It's probably possible to
recreate the original faithfully in \LaTeX if one has greater skill and more
patheince that I possess.

The dark art of indexing proved to be very challenging; entries often correspond
to parts of chapters where ideas are discussed, and even with reference to the
original text it was not always possible to guarantee that I started and ended those
blocks in the exact same places. I believe that I've been generally successful
in producing a new version of the text and supporting materials
that can be used to produce both print and electronic versions of a
foundational work in animal breeding. I apologize in advance if I've introduced
any errors during the conversion and typesetting processes.\\
~\\
Sun Prairie, Wisconsin\\
January, 2022\\
\begin{flushright}
	JOHN B. COLE
\end{flushright}